---
title: "Project1"
author: "qiankang wang"
date: "20/01/2022"
output: pdf_document
---

```{r}
knitr::opts_chunk$set(fig.width=5, fig.height=3.5) 
```


## Question a

### Implementation of SI spatial infectious disease model

```{r}
# infection function
prob_infected = function(a, b, dist) {
  return(1 - exp(-a * dist^(-b)))
}

# initialization
X = rep(1:10, 10)
Y = c()
for (i in 1:10) {
  Y = c(Y, rep(i, 10))
}

# infection one time
infection_process = function(a, b) {
  k = sample(1:100, 1)
  infected = rep(0, 100)
  infected[k] = 1
   
  dists = sqrt((X[k] - X)^2 + (Y[k] - Y)^2)
  prob = prob_infected(a, b, dists)
  idx = runif(100) <= prob
  idx[k] = FALSE
  infected[idx] = 2
  return(infected)
}

# plot the infection
plot_infection = function(a, b) {
  infected = infection_process(a, b)
  idx = infected == 0
  plot(X[idx], Y[idx], type = "p", pch = ".", xlab = "x", ylab = "y")
  points(X[!idx], Y[!idx], pch = "o", lwd = 6)
}
```

### Four realizations of this model

Big o means infection, while . means individuals not infected.

```{r}
set.seed(2022)
plot_infection(1.5, 2.0)
plot_infection(1.5, 2.0)
plot_infection(1.5, 2.0)
plot_infection(1.5, 2.0)

```

### A second version by counting individuals in set I2 and histogram of distribution of the number

```{r}
N = 100
counts = rep(0, N)
for (i in 1:N) {
  ret = infection_process(1.5, 2.0)
  # A second version to record the number by the initial infection 
  counts[i] = sum(ret == 2)
}
hist(counts, breaks = 20, main = "The number of second infection")

```

## Question b

```{r}

as = seq(1.5, 0, -0.01)
N = 1000
highest_a = NULL

set.seed(2022)
for (a in as) {
  counts = rep(0, N)
  for (i in 1:N) {
    counts[i] = sum(infection_process(a, 2.0) == 2)
  }
  if (sum(counts==0) >= 0.8*N) {
    highest_a = a
    break
  }
}

print(paste0("The highest a value is ", highest_a))
```

I tested $\alpha \in [0, 1.5]$, because smaller $\alpha$ leads lower probability of infection. Also, I made the number of total simulation equals to 1000, so I only need to the largest $\alpha$ which ensured no infection over 800 times. The largest $\alpha$ is 0.02.

